---
title: "week5-sk"
author: "Sage Kime"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male and female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap") 

## Histogram
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
    )
```

Get vectors of body_mass_g for female and male Chinstraps separately 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-test, a couple different ways

1. Is there a significant difference in body mass between male and female Chinstraps?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```



2. Are male chinstrap penguins significantly larger than females?

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")


```



# Part 2. ANOVA

## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```

Create a subset of female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex =="female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species), bins = 15) +
  facet_wrap(~species, scales = "free")

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
```


### Find summary statistics + Levene Test

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(mean_flipper = mean(flipper_length_mm),
            sd_flipper = sd(flipper_length_mm),
            sample_size = n(),
            variance = var(flipper_length_mm),
            se_flipper = sd(flipper_length_mm)/sqrt(n())
            )
penguin_f_summary


leveneTest(flipper_length_mm ~ species, data = penguin_f) 
```


### Beeswarm plot with mean +- 1sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x = species, 
                    y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary, 
             aes(x = species, y = mean_flipper),
             size = 4) +
  geom_errorbar(data = penguin_f_summary, 
                aes(x = species, 
                    ymax = mean_flipper + sd_flipper,
                    ymin = mean_flipper - sd_flipper),
                width = 0.1) +
  theme_light()
```


### One-way ANOVA + post-hoc test

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```


### Type III ANOVA w/ car package

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```

